---
title       : Visualizing adjusted linear regression
subtitle    : A shiny app
author      : Patrick Fisher
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---

## Linear regression

1. Ordinary least squares linear (OLS) regression is a commonly used statistical analysis approach for evaluating the association between two variables.
2. The association between an outcome (Y) and a predictor (X1) can be determined alone or in the context of covariates (X2, X3, etc.).
3. Inclusion/exclusion affects the estimated association between Y and X1, sometimes dramatically.
4. Visualizing the effects of including/excluding covariates can be tedious.

--- .class #id 

## Visualizing adjusted linear regression: VALR

1. Visualizaing adjusted linear regression (VALR) is a simple app that visualizes the association between two variables (Y, X1).
2. VALR allows the user to easily add or remove additional predictors from the model and re-fit the model.
3. Each time the model is re-fit, the plot updates and reports the estimated association between the selected predictor (X1) and the outcome (Y).
4. There are three parameters to be set:
  + "Brain region" - the outcome parameter
  + "Predictors" - the predictors in the model
  + "Predictor to plot" - the predictor to plot
5. Now we will walk through an example but the app can be found here
  + The app can be found here: https://fishpm.shinyapps.io/project/

--- .class #id

## VALR example

1. In the current example we are trying to predict a measure of the serotonin system in a brain area called the caudate.

2. We will first evaluate how well age predicts caudate.

3. We will then evluate how well age predicts caudate with an additional covariate (memory performance)

4. VALR provides an app for evaluating this effect.

--- .class #id

## VALR example

* Here we predict a measure in a brain area called the caudate with age.

```{r echo = F}
library(lava)
data(serotonin, package = 'lava')
serotonin$age.orig <- serotonin$age
serotonin$mem.orig <- serotonin$mem
serotonin$age <- mean(serotonin$age.orig) - serotonin$age.orig
serotonin$mem <- mean(serotonin$mem.orig) - serotonin$mem.orig

f <- 'cau ~ age'
l <- lm(f, data = serotonin)
adj <- l$residuals + l$coefficients[1]
plot(adj ~ serotonin[,'age'], pch = 19, cex = 1.2, bty = 'l', las = 1, xlab = 'Age', ylab = 'Caudate', col = 'darkblue')
abline(a = l$coefficients[1], b = l$coefficients['age'], col = 'darkblue', lwd = 3)
title(main = 'Predicting Caudate with Age', sub = 'Covariates:')
mtext(paste('Age estimate: ', signif(l$coefficients['age'],3), ', 95% CI: ', signif(confint(l)['age',1],3), '; ', signif(confint(l)['age',2],3), ', p = ', signif(summary(l)$coefficients['age',4],3), sep = ''))
```

--- .class #id

## VALR example

* Note the minor change when adding memory performance as a covariate.

```{r echo = F}
# library(lava)
# data(serotonin, package = 'lava')
# serotonin$age.orig <- serotonin$age
# serotonin$mem.orig <- serotonin$mem
# serotonin$age <- mean(serotonin$age.orig) - serotonin$age.orig
# serotonin$mem <- mean(serotonin$mem.orig) - serotonin$mem.orig

f <- 'cau ~ age + mem'
l <- lm(f, data = serotonin)
adj <- l$residuals + l$coefficients[1] + l$coefficients['mem']*l$model[,'mem']
plot(adj ~ serotonin[,'age'], pch = 19, cex = 1.2, bty = 'l', las = 1, xlab = 'Age', ylab = 'Caudate', col = 'darkblue')
abline(a = l$coefficients[1], b = l$coefficients['age'], col = 'darkblue', lwd = 3)
title(main = 'Predicting Caudate with Age', sub = 'Covariates: mem')
mtext(paste('Age estimate: ', signif(l$coefficients['age'],3), ', 95% CI: ', signif(confint(l)['age',1],3), '; ', signif(confint(l)['age',2],3), ', p = ', signif(summary(l)$coefficients['age',4],3), sep = ''))
```